\documentclass{jjsce}
%\documentclass[jscefinal]{jjsce}% 登載可決定後の最終原稿を提出するときはこちら → ページ番号なしになる

\usepackage{showkeys}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage[defaultsups]{newtxtext}
\usepackage[varg]{newtxmath}
\usepackage{bm} % 数式でボールドイタリックを使いたいとき
%% graphicx 利用の場合は以下のどちらかを選ぶ
\usepackage[dvipdfmx]{graphicx}% (u)pLaTeX + dvipdfmx の場合
%\usepackage{graphicx}% LuaLaTeXの場合
%\usepackage[dvips]{graphicx}
%\usepackage{xcolor}
\usepackage[superscript]{cite}
\usepackage{url}
\usepackage{endnotes}
\usepackage[savepos]{zref}
%% hyperref 利用の場合は以下のどちらかを選ぶ
\usepackage[dvipdfmx]{hyperref}% (u)pLaTeX + dvipdfmx の場合
\usepackage{pxjahyper}% (u)pLaTeX + dvipdfmx の場合はこれを併用する
%\usepackage[luatex,pdfencoding=auto]{hyperref}% LuaLaTeXの場合
\aboveEtitlesep20mm % 最終ページの英文タイトル部：本文末の並行止めした後のスペース調整．詳細はreadme.pdf参照．

\usepackage{jjsce-macros}

\newcommand{\fat}[1]{\boldsymbol{#1}}
\begin{document}
\jtitle{
	き裂による散乱波の反響環境下での\\時間反転集束}
%\jsubtitle{}
\etitle{
	TIME-REVERSAL FOCUSING OF ELASTODYNAMIC SCATTERED WAVES 
FROM A CRACK IN A REVERBERATING ENVIRONMENT}
%\esubtitle{}
\authorlist{%
 \authorentry{木本　和志}{Kazushi KIMOTO}{OU}
 \authorentry{斎藤　隆泰}{Takahiro SAITOH}{GU}
}
\Caffiliate[OU]{正会員　岡山大学大学院　学術研究院環境生命科学学域
（\jipcode{700--8530}岡山市北区津島中3-1-1）}{kimoto@okayama-u.ac.jp}
\affiliate[GU]{正会員　群馬大学大学院　理工学府環境創生部門
（\jipcode{376--8515}群馬県桐生市天神町1-5-1）}{t-saitoh@gunma-u.ac.jp}

%\Jbreakauthorline{4}
%\breakauthorline{4}
%\received{2022}{1}{31} Received Date を記入
%\accepted{2022}{4}{28} Accepted Date を記入
\begin{abstract}
本研究は時間反転集束の原理に基づく超音波イメージングが，多重反射が無視できない
反響環境下における超音波探傷にも有効であることを，実験と数値シミュ-ション
によって実証することを目的として行ったものである．
実験では，T継手形状を模擬した試験体を用いてスリットからの散乱波を含む
場を試験体表面で多点計測した．得られた波形は時間反転して数値波動解析モデルに
入力し，時間反転場の進展挙動を調べた．その結果，計測波形では全く検出することの
できない散乱波成分が散乱源であるスリットへ正しく集束する様子を可視化することが
できた．さらに，時間反転場をreverse time migrationの方法で画像化したところ，
散乱源位置に正しく結像することを示し，微弱な散乱波を自動的に検出して
超音波イメージングが可能であることから，時間反転集束法が継手のような
複雑形状部位の超音波探傷に有用であることを明らかにした．
\end{abstract}
\begin{Eabstract}
In this study, ultrasonic measurments and simulations are perfoemed to show 
that the ultrasonic flaw imaging based upon the time-reversal focusin concept 
works well with reverberating wavefield.
 This study demonstrates that the ultrasonic imaging based upon time-reversal 
 focusing concept works well with a reverberating wavefield using the measured 
 ultrasonic waveforms. To this end, ultrasonic measurements are performed 
 with a specimen mimkcinkg the T-shape joint geometry having a slit as an articifial flaw. 
 The measured set of ultrasonic waveformes are then time-reversed and re-emmited at 
 their observation points in the finite-difference numerical model.
 The observation on the time-reversed fields reveals that a 
 very small scattered waves from the flaw focuses back to the slit. 
 Further more reverse-time image synthesized from the forward and 
 time reversed fields can detect the scattering source successfully thus 
 demonstrates the usefulness of the time-reversal approach in the context of 
 ultrasonic flaw detectin in the reverberating environment.

time-reversal focusing fields are reconstructed numerically from measured ultrasonic signals.
In the ultrasonic measurement, reverberating field is excited in a T-shape joint specimen with and without a simulated crack. 
The measured waveforms are time-reversed and back-propagated in numerical finite-difference models of the specimen.
It is found as a result that the scattered fields back-focus successfully to the scattering sources,
and thus the wave components useful for the flaw detection may be retrieved from the reverberating waveforms.
\end{Eabstract}
\begin{keyword}
time-reversal, ultrasonic nondestructive testing, crack, scattering
\end{keyword}
\maketitle
\section{はじめに}
	\input{1_intro.tex}
\section{超音波エコー計測}
	\input{2_experiment.tex}
\section{時間反転場の集束挙動}
	\input{3_timerev.tex}
\section{RTMイメージング}
	\input{4_rtm.tex}
\section{まとめ}
本研究では，実験で計測した超音波波形を使った時間反転場の解析を行うとともに，
RTM(reverse time migration)に基づく欠陥の超音波イメージングを行った．
その結果，微弱なスリットからのエコーを，散乱源であるスリット位置に正しく
時間反転集束させられることを示した．超音波イメージングでは，
FWI(fullwave inversion)の${\rm Fr\acute{e}chet}$ Kernelに相当する
画像化関数を用い欠陥画像を合成したところ，超音波で強く照射される
スリット上半分と，スリットの起点となる切り欠き側面が正確に検出される
ことが分かった．以上のことから，時間反転法場，すなわち，動弾性問題の随伴場
の計算に基づく最適化やイメージング法は，溶接継手のような複雑な形状をした
狭い領域に対する超音波イメージングにも有用であることが示された．

時間反転場の計算結果として明らかになった散乱波の伝播経路は，
溶接ビード表面を経由し，縦波から横波へのモード変換を伴うものであった．
元々、入射波や周辺境界からの反射波や回折波に比べて微弱であることに
加えこのような複雑な経路をたどるエコーの存在を観測した超音波波形だけから
知ることは非常に困難であり，時間反転場の計算によって初めて検出
され，利用可能となるものである．
さらに，時間反転場を使ったRTMイメージングで得られた結果は，背景の
ノイズが多いということで改善の余地はあるものの，スリットとその
起点となる切り欠き側面を検出しており，同等の正確さでの画像化は
従来法である開口合成法やその類似法では経路特定と波形変化の
予想が難しいという意味で原理的に達成が困難なものである．
以上の結果は，きず周辺での回折や反射に超音波の反響が無視できず，
エコー伝播経路も複雑となり得る状況で、時間反転集束に基づく
エコー識別や画像化が，有効なアプローチとなることを示唆する．

今後は，送受信位置が時間反転場の挙動に与える影響を調べて最適化をすること，
RTMイメージングの信号-ノイズ比を改善することが課題となる．
また、超音波探傷試験への実用化に向けた課題としては，
時間反転場の計算効率を改善することが重要となる．さらに，実際の探傷には
環境振動や対象物の表面性状に影響を受けやすいレーザー振動計でなく，
圧電センサーを使った計測を行うことが望ましい。
従って，時間反転集束法に適したアレイ探触子の開発や使用も
高品質な観測データを得るために重要な課題となる。

%``\verb|\usepackage{jjsce-macros}|''
%を有効にしている場合，図表の参照に
%``\verb|\figref{}|''，
%``\verb|\tabref{}|''
%が使えます．
%\section{参考文献の引用とリスト}
%参考文献は出現順に番号を振り，その引用箇所でこのように\cite{a}上付き右括弧付き数字で指示します．
%\verb|\usepackage[superscript]{cite}|を有効にしている場合，
%``\verb|\cite{a,b,c,d}|''と記述すると，
%\cite{a,b,c,d} と記載されます．
%``文献\verb|\Cite{a}|''と記述すると、
%文献\Cite{a} と上付き添字ではなく文章中に記載されます．
\Acknowledgment % 謝辞：
「謝辞」は「結論」の後に置いて下さい．
%\appendix
%\section*{付録の見出し}
%付録のsectionが一つしかない場合は，付録に番号をつける必要がありません．
%``\verb|\section*{}|''を用いてください．
%付録のsectionが複数ある場合は，
%``\verb|\section{}|''を用いてください．

\begin{thebibliography}{9}
\bibitem{FDTD_book}
	佐藤雅弘：FDTD法による弾性振動・波動の解析入門，pp. 34-49，森北出版，2003. 
	[Satoh, M.: \textit{FDTD-hou niyoru Danseishindou\& Hadou-no-kaiseki-nyumon}, pp. 34-49, Morikita Syuppan, 2003.]
\bibitem{FDTD_KK}
	Kimoto, K. and Ichikawa, Y.:
	A finite difference method for elastic wave scattering by a planar crack with contacting faces
	, \textit{Wave Motion}., Vol. 52, pp.120-137, 2015.
\bibitem{Thomson1984}
	Thomson, R. N.:
	Transverse and longitudinal resolution of the synthetic aperture focusing technique,
	\textit{Ultrasonics}.,Vol. 22, pp.9-15, 1984. 
\bibitem{Doctor1986}
	Doctor, S. R., Hall, T. E. and Reid, L. D.:
	SAFT -the evolution of a signal processing technology for ultrasonic testing,
	\textit{ NDT International},Vol.19, No.3, pp.163-167, 1986. 
\bibitem{Schmitz2000}
	Schmitz, V.,  Chakhlov, S. and M\"{u}ller, W.:
	Experiences with synthetic aperture focusing technique in the field,
	\textit{ Ultrasonics}, Vol.38, pp.731-738, 2000.
\bibitem{Spies2012}
	Spies, M., Rieder, H., Dillh\"{o}fer, A.,  Shmitz, V. and M\"{u}ller, W.:
	Synthetic aperture focusing and time-of-flight diffraction 
	ultrasonic imaging -past and present,
	\textit{ J. Nondestruct. Eval. },Vol.31, pp.310-323, 2012.
\bibitem{a}
	本間仁，安芸皓一：物部水理学，pp. 430-463，岩波書店，1962. 
	[Honma, S. and Aki, K.: \textit{Mononobe Suiri-gaku}, pp. 430-463, Iwanami Shoten, 1962.]
\bibitem{b}
	日本道路協会：道路橋示方書・同解説IV下部構造編，pp. 110-119，1996. 
	[Japan Road Association: \textit{Dorokyo-shihosyo \& Doukaisetsu} IV Kabukouzo-hen, pp. 110-119, 1996.]
\bibitem{c}
	Shepard, F. P. and Inman, D. L.: Nearshore water circu-lation related to bottom topography and wave refraction, 
	\textit{Trans. AGU}., Vol. 31, No. 2, 1950.
\bibitem{d}
	C. R. ワイリー（富久泰明訳）：工学数学（上），pp. 123-140，ブレイン図書，1973. 
	[Wylie, C. R. (translated by Tomihisa, Y.): \textit{Advanced Eingineering Mathmatic}, Brain-tosho, 1973.]
\end{thebibliography}
\end{document}

\bibitem{Mayer1989}
	{K. Mayer, R. Marklein, K.J.Langenberg, and T. Kreutter,  
	"Three-dimensional imaging system based on Fourier transform synthetic 
	aperture focusing technique",
	{\em Ultrasonics},{\bf 28}(1990),241-255.
	}
\bibitem{Levesque2002}
	{D. L\'evesque, A. Blouin, C. N\'eron, J.-P. Monchalin, 
	"Performance of laser-ultrasonic F-SAFT imaging",
	{\em Ultrasonics},{\bf 40}(2002),1057-1063. 
	}
\bibitem{Rossignol2003}
	{C. Rossignol, H. Meri, B. Audoin, 
	"Laser generation of acoustic waves in an anisotropic medium; parametric analysis
	for application to material characterization", 
	{\em Review of Scientific Instruments}, {\bf 74}(2003), 470-473.
	}
\bibitem{Drinkwater2006}
	{ B. W. Drinkwater and P. D. Wilcox, 
	"Ultrasonic arrays for non-destructive evaluation: A review",
	{\em NDT \& E International}, {\bf 39}(2006),525-541. 
	}
\bibitem{Wilcox2007}
	{P. D. Wilcox, C. Holmes, B. W. Drinkwater,
	"Advanced reflector characterization with ultrasonic phased arrays in NDE applications",
	{\em IEEE Trans. Ultraso. Feroelect. Freq. Contr.}, {\bf 54}(8)(2007) 1541-1550.
	}
\bibitem{Holmes2008}
	{C. Holmes, B. W. Drinkwater and P. D. Wilcox,
	"Adcanced post-processing for scanned ultrasonic arrays: Application to 
	defect detection and clasification in non-destructive evaluation",
	{\em Ultrasonics},{\bf 48}(2008) 636-642.
	}
\bibitem{Shih2014}
	{ R.-C. Shih, Y.-F. Chang, C-.H. Chang, P.-Y. Tseng, 
	"Ultrasonic synthetic aperture focusing using the root-mean-square velocity"
	{\em J. Nondstruct. Eval.},{\bf 33}(2014), 12-22.
	}
\bibitem{Qin2014}
	{ K. Qin, C. Yang, F. Sun, 
	"Generalized frequency-domain synthetic aperture focusing technique for ultrasonic 
	imaging of irregularly layered objects",
	{\em IEEE Trans. Ultraso. Feroelect. Freq. Contr.}, {\bf 61}(1)(2014) 133-146.
	}
\bibitem{Stepinski2007}
	{ T. Stepinski,
	"An implementation of synthetic aperture focusing technique in frequency domain",
	{\em IEEE Trans. Ultraso. Feroelect. Freq. Contr.}, {\bf 54}(7)(2007) 1399-1408.
	}
\bibitem{Shlivinski2007}
	{A. Shlivinski and K. J. Langenberg,
	"Defect imaging with elastic waves in inhomogeneous-anisotropic materials with 
	composite geometries",
	{\em Ultrasonics},{\bf 46}(2007) 89-104.
	}
\bibitem{Schmerr1999}
	{ L. W. Schmerr, "Fundamental of ultrasonic nondestructive evaluation: 
	a modeling approach", Academic Press (1999).	
	}
\bibitem{Ceranoglu1981}
	{ A. N. Ceranoglu, and Y.-H. Pao, 
	"Propagation of elastic pulses and acoustic emission in a plate -Part 1:Theory",
	{\em J. Appl. Mech.}, {\bf 48}(1)(1981), 125-132.
	}
\bibitem{Gridin1998}
	{ D. Gridin, 
	"High-frequency asymptotic description of head waves and boundary layers surrounding the critical rays in an elastic half-space",
	{\em J. Acoust. Soc. Am.},{\bf 104}(3)(1998), 1188-1197. 
	}
\bibitem{Zhang2010}
	{ J. Zhang, B. W. Drinkwater , P. D. Wilcox, and A. J. Hunter,
	"Defect detection using ultrasonic arrays: The multi-mode total focusing method",
	{\em NDT \& E International}, {\bf 43}(2010), 123-133.
	}
\bibitem{Langenberg1989}
	{ K. J. Langenberg, "Introduction to the special issue on inverse problems",
	{\em Wave Motion}, {\bf 11}(1989),99-112.
	}
\bibitem{Kitahara2002}
	{M. Kitahara, K. Nakahata and S. Hirose,
	"Elastodynamic inversion for shape reconstruction and type classification of flaws",
	{\em Wave Motion},{\bf 36}(2002) 443-455
	}
\bibitem{Langenberg1997}
	{ K. J. Langenberg, R. B\"{a}rmann, R. Marklein, S. Irmer, H. M\"{u}ller,
	M. Brandfab, and B. Potzkai, "Electromagnetic and elastic wave scattering
	and inverse scattering applied to concrete",
	{\em NDT \& E International},{\bf 30}(4) 1997, 205-210.
	}
\bibitem{Langenberg2006}
	{ K. J. Langenberg, K. Mayer and R. Marklein, 
	" Nondestructive testing of concrete with electromagnetic and elastic waves: modeling and imaging",
	{\em Cement \& Concrete Composites}, {\bf 28}(2006), 370-383.
	}
\bibitem{DevaneyTextBook}
	{ A. J. Devaney, "Mathematical foundations of imaging, tomography and wavefield inversion",
	Cambridge University Press(2012).
	}
\bibitem{Velichko2010}
	{A. Velichko and P. D. Wilcox,
	"An analytical comparison of ultrasonic array imaging algorithm",
	{\em J. Acoust. Soc. Am.},{\bf 127}(2010), 2377-2384.
	}
\bibitem{Rose2015}
	{ L. R. F. Rose, E. Chan and C. H. Wang,
	"A comparison and extensions of algorithms for quantitative imaging
	of laminar damage in plates. I. Point spread functions and near field imaging",
	{\em Wave Motion},{\bf 58 }(2015),222-243.
	}
\bibitem{FinkTextBook}{
	M. Fink, W. A. Kuperman, J.-P. Montagner, and A.Tourin (Eds.),
	"Imaging of complex media with acoustic and seismic waves", Springer (2002)
	}
\bibitem{Prada1994}
	{ C. Prada and M. Fink, 
	"Eigenmode of the time reversal operator: A solution to selective focusing in 
	multiple-target media", 
	{\em Wave Motion}, {\bf 20}(1994), 151-163. 
	}	
\bibitem{Prada1995}
	{ C. Prada and M. Fink, 
	"Selective focusing through inhomogeneous media: the DORT method",
	{\em IEEE Ultrasonics Symposium}, (1995), 1449-1453. 
	}	
\bibitem{Mordant1999}
	{ N. Mordant, C. Prada, and M. Fink,
	"Highly resolved detection and selective focusing in a waveguide using the D.O.R.T. method",
	{\em J. Acoust. Soc. Am.},{\bf 105}(5)(1999),2634-2642. 
	}
\bibitem{Devaney2005}
	{ A. J. Devaney, E. A. Marengo, and F. K. Gruber,
	"Time-reversal-based imaging and inverse scattering of multiply scattering point targets",
	{\em J. Acoust. Soc. Am.},{\bf 118}(5)(2005),3129-3138. 
	}
\bibitem{Marengo2007}
	{ E. A. Marengo, F. K. Gruber, and F. Simonetti, 
	"Time-reversal MUSIC imaging of extended targets",
	{\em IEEE on Image Processing}, {\bf 16}(8)(2007), 1967-1984. 
	}
\bibitem{Labyed2012}
	{ Y. Labyed and L. Huang, 
	"Ultrasound time-reversal MUSIC imaging of extended targets", 
	{\em Ultrasound in Med. \& Biol.}, {\bf 38}(11)(2012), 2018-2030.
	}
\bibitem{Dominguez2005}
	{ N. Dominguez, and V. Gbiat,
	"Time domain topological gradient and time reversal analogy: 
	an inverse method for ultrasonic target detection", 
	{\em Wave Motion}, {\bf42}(1)(2005), 31-52.
	}
\bibitem{Dominguez2010}
	{ N. Dominguez, and V. Gibiat,
	"Non-destructive imaging using the time domain topological energy method",
	{\em Ultrasonics}, {\bf 50}(2010), 367-372.
	}
\bibitem{Gibiat2010}
	{ V. Gibiat and P. Shuguet,
	"Wave guide imaging through time domain topological energy",
	{\em Ultrasonics},{\bf 50}(2010), 172-179.
	}
\bibitem{Etgen2009}
	{ J. Etgen, S. H. Gray, and Y. Zhang, 
	"An overview of depth imaging in exploration geophysics",
	{\em Geophysics},{\bf 74}(6)(2009), WCA5-WCA17.
	}
\bibitem{Chung2012}
	{ W. Chung, S. Pyun, H. S. Bae, C. S.Shin, and K. J. Marfurt,
	"Implementation of elastic reverse-time migration using wavefield separation in the frequency domain",
	{\em Geophys. J. int.},{\bf 189}(2012), 1611-1625. 
	}
\bibitem{Yan2008}
	{ J. Yan and P. Sava,
	"Isotropic angle-domain elastic reverse-time migration",
	{\em Geophysics},{\bf 73}(6)(2008), S229-S239. 
	}
\bibitem{Achenbach2002}
	{ J. D. Achenbach, 
	"Modeling for quantitative non-destructive evaluation",
	{\em Ultrasonics}, {\bf 40}(2002),1-10.
	}
\bibitem{Langenberg1986}
	{K.J. Langenberg, M. Berger, Th. Kreutter, K. Mayer, and V. Schmitz,
	"Synthetic aperture focusing technique signal processing",
	{\em NDT International},{\bf 19}(3)(1986), 177-189.
	}
\bibitem{Fellinger1995}
	{ P. Fellinger, R. Marklein, K.J. Langenberg, and S. Klaholz, 
	"Numerical modeling of elastic wave propagation and scattering with EFIT 
	-Elastodynamic finite integration technique", 
	{\em Wave Motion},{\bf 21}(1995),47-66.
	}
\bibitem{Kimoto2015}
	{ K. Kimoto, and Y. Ichikawa, 
	"A finite difference method for elastic wave scattering by a planar crack 
	with contacting faces",
	{\em Wave Motion},{\bf 52}(2015),120-137.
	}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%5
\end{thebibliography}
%\input{tables.tex}
%\input{figs.tex}
%\input{caption.tex}
\end{document}

%%
%% End of file `elsarticle-template-1a-num.tex'.
\bibitem{Zhang2010}
	{J. Zhang, B. W. Drinkwater, P. D. Wilcox and A.J. Hunter,
	"Defect detection using ultrasonic arrays: The multi-mode total focusing method",
	{\em NDT \& E International}, {\bf 43}, (2010) 123-133.
	}
\bibitem{Devaney1984}
	{ A. J. Devaney and G. Beylkin, 
	"Diffraction tomography using arbitrary transmitter and receiver surfaces",
	{\em Ultrasonic Imaging},{\bf 6}(1984), 181-193.
	}
